---
title: "Assignment_2"
author: "Phill Graaf"
date: "2023-05-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Shark data tidying and plotting ####


For this section we used a Data Set from QFish.
We wanted to look at how many individuals of which species (mammal, other, shark and turtle) were found in the different locations.
We began by deleting the years that we did not need in our excel file, sorting the areas from north to south and imported the data into R:

```{r}
install.packages("tidyr", repos = "http://cran.us.r-project.org")
library("tidyr")
install.packages("ggplot2", repos = "http://cran.us.r-project.org")
library("ggplot2")
install.packages("dplyr", repos = "http://cran.us.r-project.org")
library("dplyr")
shark <- read.csv("/Users/cinematograaf/Documents/JCU/Sem 3/TIMS1/Introduction to Data Science/Homework 2/shark_data.csv")

shark
```

Because we had the variables "year" and "area" as columns, but every single species was a column itself, we wanted to make this
table "longer" since a table of a similar format with more entries would get unnecessarily wide and would be harder to analyse

We pivoted the offending columns into a new set of variables called "species" and "count":

```{r}
shark_new <- shark %>%
  pivot_longer(c(`mammal`, `other`, `shark`, `turtle`), names_to="species", values_to="count")
```

It works!
Here is our new table:

```{r}
shark_new
```
So here we have our new table!
Now we can begin plotting it:


```{r}
ggplot(data=shark_new)+
  geom_col(mapping=aes(x=factor(area), y=count, fill=species), position="dodge2")+
  scale_y_continuous(expand = c(0, 0), limits = c(0, 110), breaks = scales::pretty_breaks(n = 20))+
  scale_x_discrete(limits=shark_new$area)+
  labs(title="Relative abundance of 4 different species disaggregated by area \nfrom North to South.", x="Area", y="Number of Individuals")+
  theme_classic()+
   theme(plot.title=element_text(size=14, face="bold", hjust = 0.5), axis.title.x = element_text(size=12), axis.title.y = element_text(size=12), axis.text = element_text(size=10))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

ggsave("Shark_Plot.jpg")
```
And there we have our finished Result!