---
title: "mb5370_module_4_report"
author: "Phill Graaf"
date: "2023-05-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# import shark data

```{r}
data <- read.csv("/Users/cinematograaf/Documents/github/mb5370/Assignment 2/shark_control_program.csv")
```


#### 5.3 What is a tibble? ####

install.packages("tidyverse")

library(tidyverse)

vignette("tibble")


iris # look at iris

str(iris) # check it out - what type is it? how many rows?

as_tibble(iris)

# building a tibble from scratch:

tibble(
  x=1:5,
  y=1,
  z=x^2+y # call new variables to produce new column values!
)

# R will not let you do this using "data.frame"

# how would you build this data frame using the old "data.frame" function?

x=c(1,2,3,4,5)
y=1
data.frame(c(
    x=1:5,
    y=1,
    z=x^2+y))
))


# non-sematic names

tb <- tibble(
`:)` = "smile",
` ` = "space",
`2000` = "number"
)

tb

# tribbles

tribble(
  ~x, ~y, ~z,
  #--|--|----
  "a", 2, 3.6,
  "b", 1, 8.5
  )

as_tibble(iris) # only shows the first 10 rows

tibble(
  a=lubridate::now()+runif(1e3)*86400,
  b=lubridate::today()+runif(1e3)*30,
  c=1:1e3,
  d=runif(1e3),
  e=sample(letters, 1e3, replace=TRUE)
)

# this tibble tells us we have a data-time column (dttm),a date column, an integer, a doublem and a character



# nycflights ####

install.packages("nycflights13")
library(nycflights13)

nycflights13::flights %>%
  print(n=10, width=Inf)

# setting global options for your R session

options(tibble.width=inf)



# useful things
# pulling out a signle variable

df <- tibble(
  x=runif(5),
  y=rnorm(5)
  )

# extract by name
df$x

# same same but different:

df[["x"]]

# extract by row position
df[[1]]

# extract by exact position
df[[2,2]]



df %>% .$x

df %>% .[["x"]]


# tibbles will warn you if a variable doesn't match exactly what's in the dataframe

df <- tibble(
  xxx=runif(5),
  y=rnorm(5)
  )

df$xx # warning message





#### shark data tidying and plotting ####


# for this section we used the "shark ???? Data Sert from Fish????
# we wanted to look at how many individuals of which species (mammal, other, shark and turlte) were found in the different locations
# we began by deleting the years that we did not need in our excell file and imported the data into R:

```{r}
shark <- read.csv("/Users/cinematograaf/Documents/JCU/Sem 3/TIMS1/Introduction to Data Science/Homework 2/shark_data.csv")

shark
```

# because we had the variables "year" and "area" as columns, but every single species was a column itself, we wanted to make this
# table "longer" since a table of a similar format with more entries would get unnecessarily wide and would be harder to analyse

# we pivoted the offending columns into a new set of variables called "species" and "count":

```{r}
shark_new <- shark %>%
  pivot_longer(c(`mammal`, `other`, `shark`, `turtle`), names_to="species", values_to="count")
```

# it works!
# here is our new table

```{r}
shark_new
```
# so here we have our new table
# now we can begin plotting it


```{r}
ggplot(data=shark_new)+
  geom_col(mapping=aes(x=factor(area), y=count, fill=species), position="dodge2")+
  scale_y_continuous(expand = c(0, 0), limits = c(0, 110), breaks = scales::pretty_breaks(n = 20))+
  scale_x_discrete(limits=shark_new$area)+
  labs(title="Relative abundance of 4 different species disaggregated by area \nfrom North to South.", x="Area", y="Number of Individuals")+
  theme_classic()+
   theme(plot.title=element_text(size=14, face="bold", hjust = 0.5), axis.title.x = element_text(size=12), axis.title.y = element_text(size=12), axis.text = element_text(size=10))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

ggsave("Shark_Plot.jpg")


- ordering a factor!
- xlim to put x-axis to 0!


# this works!


shark_plot <- shark_new %>%. 
  ggplot(aes(x=area, y=count)+
  geom_col(fill=species)

shark_plot
```





#### 6.0 Spatial data in R ####

```{r}
install.packages("sf")
install.packages("terra")
install.packages("tmap")

# load into R library!
library("tidyverse") 
library("sf") # simple features
library("terra") # for raster
library("tmap") # thematic maps are geographical maps in which spatial data distributions are visualised 
```
# first we will load the copepod data in R

```{r}
library(readr)
dat <- read_csv("/Users/cinematograaf/Documents/github/mb5370/spatial data/data-for-course/copepods_raw.csv")
dat
```
# silk_id is the ID for each of the silks, onto which plankton are recorded


# now we plot the coordinates for the samples (segments of the CPR silks)

```{r}
library(ggplot2)
ggplot(dat)+
  aes(x=longitude, y=latitude, colour=richness_raw)+
  geom_point()
```
# this looks good but it is not a map!
# now let's look at the richness data, our main variable for analysis
```{r}
ggplot(dat, aes(x=latitude, y=richness_raw))+
  stat_smooth()+
  geom_point()
```
# we will not repeat the above map for richness

```{r}
sdat <- st_as_sf(dat, coords=c("longitude", "latitude"),
                 crs=4326)
```
# "st_as_sf" converts different data types to simple features
# "dat" is our original data
# "coords" gives the names of the columns that relate to the spatial coordinates (in order of X coordinate followed by Y  coordinate)
# "crs" stands for coordinate reference system which we will discuss next


#### 6.9 Coordinate reference systems ####

```{r}
crs4326 <- st_crs(4326) 
crs4326 # look at the whole CRS
crs4326$Name # pull out just the name of the crs
```
#now let's check out what the code looks like
```{r}
crs4326$wkt # crs in well-known text format
```
#### 6.10 Feature collection (points) ####
# Let's now look at what we created with sdat
```{r}
sdat
```
#### 6.11 Cartography ####
# now let's get into mapping! "sf" has simple plotting features like:
```{r}
plot(sdat["richness_raw"]) # selects a single variable: richness
```
```{r}
plot(sdat) # creates a panel for every single variable in our data frame
```
#### 6.12 Thematic maps for communication

```{r}
# using tmap
tm1 <- tm_shape(sdat)+
  tm_dots(col="richness_raw")
```
# "tm_dots" to plot dots of the coordinates (other options are tm_polygons)
# "tm_symbols" and many oithers we'll see later
# we've chosen "richness_raw" as the colour scale

# using a different colour ramp:
```{r}
tmap_save(tm1, filename="Richness-map.png",
          width=600, height=600)
```
#### 6.13 Mapping spatial polygons as layers ####

# 6.12.1 Loading shapefiles

```{r}
aus <- st_read("/Users/cinematograaf/Documents/github/mb5370/spatial data/data-for-course/spatial-data/Aussie/Aussie.shp")

shelf <- st_read("/Users/cinematograaf/Documents/github/mb5370/spatial data/data-for-course/spatial-data/aus_shelf/aus_shelf.shp")

aus
```
# 6.13.2 mapping your polygons
# tmap makes it very straightforward to make a map of polygons
```{r}
tm_shape(shelf)+
  tm_polygons()
```
# we have indicated the shape of our map (shelf) and have added a command "bbox=sdat" to expand the extent of the map so it depicts all of our copepod data points
# we then added the shape of australia (aus) on top of the shelf and finally our copepod data (sdat) in the form of points using "tm_dots()"

```{r}
tm_shape(shelf, bbox=sdat)+
  tm_polygons()+
  tm_shape(aus)+
  tm_polygons()+
  tm_shape(sdat)+
  tm_dots()  
```
# 6.14 Exploring t_map
# changing map style to "beaver"

```{r}
tmap_style("beaver")


tm_shape(rsst)+
  tm_raster(palette="-RdBu", title="SST")+
  tm_shape(shelf, bbox=sdat)+
  tm_polygons(col="aquamarine",title="cont. shelf")+
  tm_shape(aus)+
  tm_polygons(col="adm", midpoint=0, title="State")+
  tm_shape(sdat)+
  tm_dots(col="vessel", title="Vessel")+
  tm_compass()

```
```{r}
rsst <- rast("/Users/cinematograaf/Documents/github/mb5370/spatial data/data-for-course/spatial-data/MeanAVHRRSST/MeanAVHRRSST.grd")
plot(rsst)
tm_shape(rsst)+
  tm_raster(palette="-RdBu", title="SST")+
  tm_shape(sdat)+
  tm_dots(col="richness_raw",
          palette="Greys",
          title="Species #")+
  tm_compass()

tmap_save()
```


```{r}
vignette("tmap-getstarted")
```


```{r}
tmap_save()
?tmap_save()
```

